### 2016-11-20 paper/supplemental/architecture.tex
Add outline for supplemental while we are waiting for CREMI results.

\section{Classifier}

\subsection{CNN Architecture}
2 vs 4 layers vs resnet

\subsection{Training Parameters}

\subsection{Automatic Method Threshold $p_t$}