### 2016-11-20 paper/supplemental/userstudy.tex

\section{Forced Choice User Experiment}

\subsection{Recruitment and Participation}

flyers, participant list

\subsection{Example Classifications}

Where did users make a mistake?

\subsection{Subjective Responses}

NASA TLX ANOVA analysis