### 2016-11-22 paper/supplemental/leftcylinder.tex

\section{L. Cylinder Results}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cyltrails.pdf}
\caption{Performance comparison of Plaza's focused proofreading and our guided proofreading on the L. Cylinder dataset. All measurements are reported as median VI, the lower the better. We compare automatic selection with threshold and the selection oracle for accepting or rejecting corrections using each method. In all cases, guided proofreading yields better results with fewer corrections.}
\label{fig:cylboxplot}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=.7\linewidth]{gfx/cylboxplot.pdf}
\caption{VI distributions of guided proofreading (GP) and focused proofreading (FP) output across slices of the L. Cylinder dataset, with different error correction approaches. The variation resulting from performance of FP with automatic selection is $7.8\times$ higher than GP (as indicated by the arrow), with median VI of $2.75$ and $SD=0.789$.}
\label{fig:cylboxplot}
\end{figure}