### 2017-03-27 paper/supplemental/cremi.tex
Update plots for the supplemental material.

\section{Additional Experiments}
\label{sec:addexp}

\paragraph{CREMI A/B/C.} As part of the MICCAI 2016 challenge on circuit reconstruction from electron microscopy images (CREMI), six ssTEM datasets were made publicly available\footnote{\scriptsize{\url{http://www.cremi.org}}},  each $1250\times1250\times125$ voxels. Since only three datasets include manually-labeled `ground truth', we use these three volumes for our experiments. The volumes are part of an adult fruit fly (Drosophila melanogaster) brain. The resolution of all three datasets is $4\times4\times40~\text{nm}^3\text{/voxel}$. We evaluate error detection and correction on subvolumes of CREMI A/B/C with the dimensions $1250\times1250\times5$ voxels. The subvolumes were cut from the last 25 sections of each of the three datasets and unseen during training. We compare focused proofreading and guided proofreading with automatic selection ($p_t=0.95$) and selection oracle.

\paragraph{Retraining.} Since the CREMI data is a different species, we simply retrain our split error classifier as well as focused proofreading by Plaza~\cite{focused_proofreading}. For this, we use the first 100 sections of each of the three CREMI datasets combined as training data. All parameters are unchanged and left as reported in the paper.

%\paragraph{Error correction.} In all three datasets merge error detection found over 300 merge errors. Unfortunately, merge error correction crashed because of a software error on our part. Therefore, we only evaluate split error detection and correction on subvolumes of CREMI A/B/C with the dimensions $1250\times1250\times5$ voxels. The subvolumes were cut from the last 25 sections of each of the three datasets and unseen during training. We compare focused proofreading and guided proofreading with automatic selection ($p_t=0.95$) and selection oracle.

\subsection{CREMI A}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiA_trails.pdf}
\caption{Performance comparison of Plaza's focused proofreading and our guided proofreading on 5 sections of the CREMI A dataset. All measurements are reported as median VI, the lower the better. The threshold for automatic selection is $p_t=0.95$ (green line). The slope of the selection oracle shows that guided proofreading reduces VI faster.}
\label{fig:cremiAtrails}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiAboxplot.pdf}
\caption{VI distributions of guided proofreading (GP) and focused proofreading (FP) output across slices of the CREMI A dataset, with different error correction approaches. The variation resulting from performance of FP with automatic selection is $5.4\times$ higher than GP (as indicated by the arrow), with median VI of $5.32$ and $SD=0.009$. GP does not reach the best possible VI as discussed in the text.}
\label{fig:cremiAboxplot}
\end{figure}

Figure~\ref{fig:cremiAtrails} and~\ref{fig:cremiAboxplot} compare Plaza's focused proofreading and guided proofreading on the 5 sections of the CREMI A dataset.

\paragraph{Selection oracle.} With focused proofreading, the selection oracle reduces median VI to 0.928, $SD=0.043$ from an initial median VI of 1.06 ($SD=0.055$). 532 corrections out of 3707 were accepted. Guided proofreading does not reach the best possible VI, however, reduces VI faster with less corrections to 0.941 ($SD=0.04$). Out of 4463 corrections, 1275 were accepted.

\paragraph{Automatic selection with threshold.} Not surprisingly, focused proofreading performs poorly when ran automatically (VI of 5.32, $SD=0.009$). Guided proofreading is able to reduce VI to 0.989 ($SD=0.043$) with $p_t=0.95$.

\subsection{CREMI B}

Figure~\ref{fig:cremiBtrails} and~\ref{fig:cremiBboxplot} show the results on the CREMI B dataset.

\paragraph{Selection oracle.} Focused proofreading is able to reduce median VI to 1.29, $SD=0.031$ from an initial median VI of 1.63 ($SD=0.025$). Out of 1959 corrections, the selection oracle accepted 517. With guided proofreading, the median VI is reduced to 1.30, $SD=0.03$ while accepting 1111 corrections out of 3073.

\paragraph{Automatic selection with threshold.} Focused proofreading results in a VI of 4.25 ($SD=0.07$). Guided proofreading reduces median VI to 1.43 ($SD=0.038$).

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiB_trails.pdf}
\caption{Split error correction by Plaza's focused proofreading and our guided proofreading compared on the CREMI B dataset. All measurements are reported as median VI, the lower the better. Automatic selection with threshold (green line) yields reasonable performance using guided proofreading.}
\label{fig:cremiBtrails}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiBboxplot.pdf}
\caption{VI distributions of guided proofreading (GP) and focused proofreading (FP) output across 5 sections of the CREMI B dataset. We compare automatic selection and oracle selection. The variation resulting from performance of FP with automatic selection is $3\times$ higher than GP (as indicated by the arrow), with median VI of $4.25$ and $SD=0.07$.}
\label{fig:cremiBboxplot}
\end{figure}

\subsection{CREMI C}

The results of split error correction using focused proofreading and guided proofreading on the CREMI C subvolume are shown in Figure~\ref{fig:cremiCtrails} and~\ref{fig:cremiCboxplot}.

\paragraph{Selection oracle.} With focused proofreading, the initial median VI of 1.75 ($SD=0.086$) is reduced to 1.45 ($SD=0.056$) with 670 accepted corrections out of 2694. Guided proofreading is able to reduce the VI to 1.47 ($SD=0.06$). Here, the oracle accepted 1531 out of 4332 corrections.

\paragraph{Automatic selection with threshold.} Focused proofreading results in a VI of 4.81 ($SD=0.03$). Guided proofreading with $p_t=0.95$ reduces median VI to 1.57 ($SD=0.081$).

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiC_trails.pdf}
\caption{Performance comparison of Plaza's focused proofreading and our guided proofreading on the CREMI C dataset. Lower VI scores are better. Guided proofreading corrects the initial segmentation faster with less corrections than focused proofreading. The green line shows the automatic threshold $p_t=0.95$.}
\label{fig:cremiCtrails}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/cremiCboxplot.pdf}
\caption{VI distributions of guided proofreading (GP) and focused proofreading (FP) output across the CREMI C subvolume, with different error correction approaches. The variation resulting from performance of FP with automatic selection is $3\times$ higher than GP (as indicated by the arrow), with median VI of $4.81$ and $SD=0.08$.}
\label{fig:cremiCboxplot}
\end{figure}