### 2017-04-04 paper/arxiv/conclusion.tex
Latest updates including supplemental for submission and the arXiv versions.

\section{Conclusions}

Humans are the bottleneck when proofreading segmentation data and minimizing the manual labor is the goal. Our classifiers suggest potential errors and corrections better than existing methods. This reduces the time spent finding and correcting errors.
Our experiments also show that automatic proofreading has potential to further reduce human involvement. This will be the target of future research. We provide our framework and data as free and open research at \url{http://rhoana.org/guidedproofreading/}.

\section*{Acknowledgements}
We would like to thank Stephen Plaza for detailed explanations of focused proofreading and Toufiq Parag for the configuration of the NeuroProof classifier.