### 2017-04-04 paper/arxiv/guidedproofreading_review.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

%\usepackage{cvpr}
\usepackage{iccv}
\usepackage{microtype}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{authblk}
% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=blue,citecolor=blue,bookmarks=false]{hyperref}

\newcommand{\JT}[1]{\textcolor{blue}{JT: #1}}
\newcommand{\HP}[1]{\textcolor{red}{HP: #1}}
\newcommand{\DFH}[1]{\textcolor{red}{DH: #1}}

\iccvfinalcopy % *** Uncomment this line for the final submission

%\def\cvprPaperID{1234} % *** Enter the CVPR Paper ID here
\def\iccvPaperID{1234}
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\ificcvfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Guided Proofreading of Automatic Segmentations for Connectomics}

\author[1,2]{Daniel Haehn\thanks{Corresponding author,  \url{haehn@seas.harvard.edu}}}
\author[1,2]{Verena Kaynig}
\author[3]{James Tompkin}
\author[2]{Jeff W. Lichtman}
\author[1,2]{Hanspeter Pfister}
\affil[1]{Harvard Paulson School of Engineering and Applied Sciences, Cambridge, MA 02138, USA}
\affil[2]{Harvard Brain Science Center, Cambridge, MA 02138, USA}
\affil[3]{Brown University, Providence, RI 02912, USA}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
%
Automatic cell image segmentation methods in connectomics produce merge and
split errors, which require correction through proofreading. Previous research
has identified the visual search for these errors as the bottleneck in
interactive proofreading. To aid error correction, we develop two classifiers
that automatically recommend candidate merges and splits to the user. These
classifiers use a convolutional neural network (CNN) that has been trained with
errors in automatic segmentations against expert-labeled ground truth. Our
classifiers detect potentially-erroneous regions by considering a large context
region around a segmentation boundary. Corrections can then be performed by a
user with yes/no decisions, which reduces variation of information $7.5\times$ faster than previous
proofreading methods. We also present a fully-automatic mode that uses a
probability threshold to make merge/split decisions. Extensive experiments using
the automatic approach and comparing performance of novice and expert users
demonstrate that our method performs favorably against state-of-the-art
proofreading methods on different connectomics datasets.
%
\end{abstract}

%%%%%%%%% BODY TEXT
\input{introduction}

\input{related_work}

\input{method}

\input{evaluation}

\input{results}

\input{conclusion}

{\small
\bibliographystyle{ieee}
\bibliography{connectomics}
}

\end{document}