### 2017-04-04 paper/supplemental/guidedproofreading_supplemental_arxiv.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

%\usepackage{cvpr}
\usepackage{iccv}
\usepackage{microtype}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{floatrow}
\usepackage{makecell}
\usepackage{authblk}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\newfloatcommand{capbtabbox}{table}[][\FBwidth]
% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,linkcolor=blue,citecolor=blue,bookmarks=false]{hyperref}

\newcommand{\JT}[1]{\textcolor{blue}{JT: #1}}
\newcommand{\HP}[1]{\textcolor{red}{HP: #1}}
\newcommand{\DFH}[1]{\textcolor{red}{DH: #1}}

\iccvfinalcopy % *** Uncomment this line for the final submission

%\def\cvprPaperID{0947} % *** Enter the CVPR Paper ID here
\def\iccvPaperID{0922}
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\ificcvfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Guided Proofreading of Automatic Segmentations for Connectomics\\~\\\textit{Supplemental Material}}

\author[1,2]{Daniel Haehn\thanks{Corresponding author,  \url{haehn@seas.harvard.edu}}}
\author[1,2]{Verena Kaynig}
\author[3]{James Tompkin}
\author[2]{Jeff W. Lichtman}
\author[1,2]{Hanspeter Pfister}
\affil[1]{Harvard Paulson School of Engineering and Applied Sciences, Cambridge, MA 02138, USA}
\affil[2]{Harvard Brain Science Center, Cambridge, MA 02138, USA}
\affil[3]{Brown University, Providence, RI 02912, USA}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT

%%%%%%%%% BODY TEXT

\input{architecture}
\input{leftcylinder}
\input{cremi}
\input{userstudy}

\newpage
{\small
\bibliographystyle{ieee}
\bibliography{connectomics}
}

\end{document}