### 2018-03-08 paper/supplemental/architecture.tex

\section{Classifier}

\subsection{Architecture}
We explored different architectures for the convolutional neural network (CNN) for split error detection. We compare traditional CNN architectures versus residual networks~\cite{resnet} (Tab.~\ref{tab:architecture}). The traditional architecture with dropout regularization generalized better than residual networks on unseen testing data.

\begin{table}[t]
\caption{Traditional CNN Architecture versus Residual Network Architecture \cite{resnet}. All configurations are compared using the same parameters. Our final choice (indicated by *) trains relatively fast and performs better.}%While the training of our classifier is more expensive, testing accuracy is superior. }

\small{
\begin{tabular}{@{}ll|l@{}}
	\toprule
     ~ & \textbf{Traditional Network} & \textbf{Residual Network}  \\ \midrule	
\begin{tabular}{@{}r|@{}}
Conv. Layers \\
Dropout Reg. \\
Cost [m] \\
Test. Acc. \\
Prec./Recall \\
F1 Score\\
~
\end{tabular} & 
\begin{tabular}{@{}c|c@{}}
 2 & 4 \\
 y & y \\
 27.5 & 383 \\
 0.925 & 0.94 \\
 0.93/0.93 &
 0.94/0.94 \\
 0.93 & 0.94 \\
 ~&*
\end{tabular}
& 
\begin{tabular}{@{}c|c@{}}
 5 & 13 \\
 y & n \\
5080 & 1094 \\
 0.93 & 0.90 \\
 0.7/0.53 &
 0.74/0.66 \\
 0.39 & 0.64\\
 ~&~

\end{tabular}

\end{tabular}
\hspace{2mm}
\hrule
}
\label{tab:architecture}
\end{table}

\subsection{Training Parameters}

We performed a limited brute force parameter search to tune the split error classifier (Tab.~\ref{tab:parametersearch}). This resulted in 3240 different CNN configurations which were evaluated on 10\% of our training data. Learning rate and momentum ranges are defined linearly across 2000 epochs.

\begin{table}[t]
\caption{Brute force parameter search for the split error classifier. The final parameters are highlighted.}%While the training of our classifier is more expensive, testing accuracy is superior. }

\small{
\begin{tabular}{@{}ll}
	\toprule
     \textbf{Parameter} & \textbf{Search Space}  \\ \midrule	
\begin{tabular}{@{}l@{}}
Filter size: \\
No. Filters 1: \\
No. Filters 2-4: \\
Dense units:\\
Learning rate: \\
Momentum: \\
Mini-Batchsize: 
\end{tabular} & 
\begin{tabular}{@{}l@{}}
\textbf{3x3}, 5x5, 9x9, 13x13\\
32, 48, \textbf{64} \\
32, \textbf{48}, 64 \\
256, \textbf{512} \\
0.00001, 0.0001, 0.001, 0.01, \textbf{0.03-0.00001} \\
0.9, 0.95, \textbf{0.9-0.999} \\
10, 100, \textbf{128}
\end{tabular}

\end{tabular}
\hspace{2mm}
\hrule
}
\label{tab:parametersearch}
\end{table}

\subsection{Automatic Method Threshold $p_t$}
For automatic selection, we observed a threshold $p_t=0.95$ as stable when evaluating on previously unseen testing data (Mouse S1 AC3 Open Connectome Project dataset). This means that automatic selection stops once all borders with $p_t\geq0.95$ were proofread. Figure~\ref{fig:thresh} shows split error classification on three randomly selected subvolumes ($700\times700\times2$ voxels) of AC3. In all cases, the threshold $p_t=0.95$ reduces VI.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{gfx/ptplot.pdf}
\caption{Observations of probability thresholds $p_t$ during automatic selection on three different subvolumes of previously unseen testing data. The dashed lines show when $p_t=0.95$ is reached.}
\label{fig:thresh}
\end{figure}

\subsection{Input channel contributions}

All four input channels help to reduce VI (Table~\ref{tab:input_channels}). As identified by Bogovich~\etal~\cite{BogovicHJ13}, image data adds intracellular structures (e.g., vesicles) to the decision process, and membrane probabilities include global knowledge of the staining protocol to highlight cell membranes. Then, the label channel provides knowledge about neuron shapes while the dilated mask of the border covers the gap of extra-cellular space. Adding the dilated mask of the border decreases VI. %\JT{From the table, it might make sense to test just Label+Border, because the other two decrease VI...}

\begin{table}[h]
\caption{Automatic selection on the AC4 subvolume with $p_t=0.95$ using the guided proofreading classifier; median VI reduction in ascending order. The combination of all four input channels performs best.}
\resizebox{.9\linewidth}{!}{
\begin{tabular}{lrrrr}
\toprule
\makecell{Input channels} & \makecell{VI reduction} \\
\midrule
%\emph{Prob. + Border (\textbf{R4} request)} & -0.080\\
%\emph{Prob. + Label + Border (\textbf{R3} request)} &  TBA\\
%\emph{Label + Border} & -0.008\\
%\emph{Image + Prob.} &  -0.094\\
%\emph{Image + Prob. + Border} & -0.045\\
%\emph{Image + Prob. + Label} & 0.038\\
%\emph{Image + Prob. + Label + Border} & 0.065\\

\emph{Image + Prob.} &  -0.094\\
\emph{Prob. + Border} & -0.080\\
\emph{Image + Prob. + Border} & -0.045\\
\emph{Label + Border} & -0.008\\
\emph{Image + Prob. + Label} & 0.038\\
\emph{Prob. + Label + Border} &  0.041\\
\emph{Image + Prob. + Label + Border} & 0.065\\
\bottomrule
\end{tabular} 
}
\label{tab:input_channels}
\end{table}

\subsection{Merge Error Detection Pseudo Code}

We provide pseudo code on how we detect merge errors to foster understanding of the reported algorithm (Alg.~\ref{code:mergeerror}). In our experiments, we use $N=50$ iterations.

\begin{algorithm}[H]
\caption{Merge Error Detection for a label \emph{l}}\label{code:mergeerror}
\begin{algorithmic}[1]
%\State{mergeErrors = []}
%\For{\emph{l} in $\forall$labels}
	\State{$l_d$ = \textbf{dilate}(\emph{l}, 20)}
	\State{invImage = \textbf{invert}(image)}
	\For{N iterations}
		\State{$s_1$, $s_2$ = \textbf{randomSeedsOnBoundary}($l_d$)}
		\State{wsImage = \textbf{watershed}(invImage, $l_d$, $s_1$, $s_2$)}
		\State{border = \textbf{border}(wsImage)}
		\State{$p$ = \textbf{rank}(border)}
		\State{$p_\text{merge}$ = $1-p$}	
	\EndFor
	\State{\textbf{find}(max $p_\text{merge})$}

\end{algorithmic}
\end{algorithm}

\subsection{Limitations}
Guided proofreading works on 2D image sections. This enables error correction without a computationally expensive alignment process. However, the output requires an additional (block-)merging step prior to 3D analysis. Several software packages exist for this purpose.

As described in Section~\ref{sec:addexp}, the guided proofreading classifier has to be retrained if used on a different species than mouse. In our experiments, parameters did not need to be changed.

\section{Automatic Segmentation Pipeline}

We create a dense automatic segmentation of electron microscopy data using a combination of a U-net~\cite{RonnebergerFB15} and the GALA agglomeration method~\cite{nunez2014graph}. To not bias, these classifiers are trained on different data than GP (Tab.~\ref{tab:trainingdata}).

\begin{table}[h]
\caption{Training data of membrane detection (U-Net / GALA) vs.~training data of GP vs.~test data.}%While the training of our classifier is more expensive, testing accuracy is superior. }
\vspace{-1mm}
\resizebox{\linewidth}{!}{
\begin{tabular}{lrrrr}
\toprule
%Dataset & \makecell{Training Set\\Membrane Detection (U-Net)} & \makecell{Training Set\\Guided Proofreading} \\ 
\makecell{Training Set\\U-Net / GALA} & \makecell{Training Set\\GP} & \makecell{Test Set\\GP}\\ 
\midrule
\makecell{AC3+AC4\\($1024\times1024\times175$vx)} & \makecell{L. Cylinder\\($2048\times2048\times250$vx)} &
\makecell{$\text{L. Cylinder}_\text{test}$\\($2048\times2048\times50$vx)}
\\ 
\makecell{AC4 excl. test\\ ($1024\times1024\times90$vx)} &  \makecell{L. Cylinder\\($2048\times2048\times250$vx)} &
\makecell{$\text{AC4}_\text{test}$ subvolume\\($400\times400\times10$vx)} \\ 
\makecell{AC3+AC4\\($1024\times1024\times175$vx)} & \makecell{CREMI A/B/C\\($1250\times1250\times300$vx)} &
\makecell{$\text{CREMI A/B/C}_\text{test}$\\($1250\times1250\times15$vx)} \\ 
\bottomrule
\end{tabular} 
}
\label{tab:trainingdata}
\end{table}

GALA uses a random forest classifier to agglomerate segments. We use an agglomeration level of 0.3 (after a grid search). We follow the method by Knowles-Barley~\etal as described in~\cite{rhoananet}.