### 2018-03-08 paper/conclusion.tex
Update acknowledgements.

\vspace{-2mm}
\section{Conclusions}

%The task of automatic cell boundary segmentation is difficult, and trying to improve such segmentations automatically as a post-process through merge and split error correction is, in principle, no different than trying to improve the underlying cell boundary segmentation. Due to the task difficulty, manual proofreading of connectomics segmentations is necessary, but it is a time consuming and error-prone task. Humans are the bottleneck and minimizing the manual labor is the goal.
%We have trained classifiers to find merge and split errors in automatic segmentations of connectomics data. We then reduce the correction of these errors to simple yes/no decisions. This allows proofreading with better performance than existing systems.
%Our experiments also show that automatic proofreading works. This has the potential to further reduce the manual labor and will be the target of future research. To encourage testing of our proposed architecture and replication of our experiments, we provide our framework and data as free and open research at (link omitted for review).
%Minimizing manual labor is the goal when proofreading large automatic segmentations. Our classifiers suggest potential errors and corrections better than existing systems. This is especially important for novice users. Novices report difficulties when using fully interactive software but are invaluable to proofread the shear amounts of connectomics data.
%Our experiments also show that automatic proofreading works. This has the potential to further reduce human involvement and will be the target of future research. We provide our framework and data as free and open research at (link omitted for review).

Manual proofreading of connectomics segmentations is necessary, but it is a time consuming and error-prone task. We attempt to minimize manual labor and reduce the human bottleneck. Our classifiers can detect errors and suggest potential corrections to automatically improve segmentations, but performance can increase with guided human attention. Our approach makes it easier for both novices and experts to proofread successfully, and is significantly better than the existing baseline at this task. This helps reduce the human time spent finding and correcting errors, and so makes it faster to generate new ground truth label data. We provide our framework and data as free and open research\footnote{\scriptsize{\url{http://rhoana.org/guidedproofreading/}}}.

%Our experiments also show that automatic proofreading has potential to further reduce human involvement. We provide our framework and data as free and open research at (link omitted for review).%\url{http://rhoana.org/guidedproofreading/}.

\vspace{-1mm}
\section*{Acknowledgements}
%We would like to thank Stephen Plaza for detailed explanations of \emph{Focused proofreading} and Toufiq Parag for the configuration of the \emph{NeuroProof} classifier. This research was supported in part by NSF grants IIS-1447344 and IIS-1607800, and by the Intelligence Advanced Research Projects Activity (IARPA) via Department of Interior/Interior Business Center (DoI/IBC) contract number D16PC00002.
We would like to thank Stephen Plaza for detailed explanations of \emph{Focused proofreading} and Toufiq Parag for the configuration of the \emph{NeuroProof} classifier. This research was supported in part by NSF grants IIS-1447344 and IIS-1607800, and by IARPA contract D16PC00002.